---
title: "Reproduction of Table 1"
output: github_document
---

<!-- Table1.md is generated from Table1.Rmd Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
```

This document reproduces the simulation results reported in Table 1 of the
paper.

> Ciardulli, S., Fontana, N., Vantini, S., & Ieva, F. (2026). Generalized
> propensity score weighting for functional causal inference framework.
> https://arxiv.org/abs/2608.03200

## Simulation data

For each combination of the treatment–covariate and covariate–outcome
relationships, we generate 200 independent data sets, each with a sample
size of 200. The same simulated data sets are reused across the different
weighting and PVE specifications within each data-generating scenario.

```{r settings}
library(tidyverse)
box::use(./R/data_generation)
box::use(estimate = ./R/estimate_causal_effect)

simulation_settings <- tribble(
  ~ `Treat-Cov`, ~ `Cov-Outcome`, ~ `PVE`, ~ `PVE*`, ~ `Weight`, 
    "Linear",      "Linear",        "",      "0.95",   "Unweighted", 
    "Linear",      "Linear",        "",      "0.99",   "Unweighted", 
    "Linear",      "Linear",        "0.95",  "0.95",   "Weighted",
    "Linear",      "Linear",        "0.95",  "0.99",   "Weighted",
    "Linear",      "Linear",        "0.99",  "0.95",   "Weighted",
    "Linear",      "Linear",        "0.99",  "0.99",   "Weighted",
    "Non-linear",  "Linear",        "",      "0.95",   "Unweighted", 
    "Non-linear",  "Linear",        "",      "0.99",   "Unweighted", 
    "Non-linear",  "Linear",        "0.95",  "0.95",   "Weighted",
    "Non-linear",  "Linear",        "0.95",  "0.99",   "Weighted",
    "Non-linear",  "Linear",        "0.99",  "0.95",   "Weighted",
    "Non-linear",  "Linear",        "0.99",  "0.99",   "Weighted",
    "Linear",      "Non-linear",    "",      "0.95",   "Unweighted", 
    "Linear",      "Non-linear",    "",      "0.99",   "Unweighted", 
    "Linear",      "Non-linear",    "0.95",  "0.95",   "Weighted",
    "Linear",      "Non-linear",    "0.95",  "0.99",   "Weighted",
    "Linear",      "Non-linear",    "0.99",  "0.95",   "Weighted",
    "Linear",      "Non-linear",    "0.99",  "0.99",   "Weighted",
    "Non-linear",  "Non-linear",    "",      "0.95",   "Unweighted", 
    "Non-linear",  "Non-linear",    "",      "0.99",   "Unweighted", 
    "Non-linear",  "Non-linear",    "0.95",  "0.95",   "Weighted",
    "Non-linear",  "Non-linear",    "0.95",  "0.99",   "Weighted",
    "Non-linear",  "Non-linear",    "0.99",  "0.95",   "Weighted",
    "Non-linear",  "Non-linear",    "0.99",  "0.99",   "Weighted",
)

n_replications <- 200 # R in the paper
sample_size <- 200    # n in the paper

get_data_impl <- function(treatment_covariate_relation, 
                          covariate_outcome_relation) {
  map(
    seq_len(n_replications), 
    \(dummy) data_generation$simulate_data(
      sample_size, treatment_covariate_relation, covariate_outcome_relation)
  )
}
get_data <- memoise::memoise(get_data_impl)

set.seed(314)
list_settings <- simulation_settings |> split(seq_len(nrow(simulation_settings)))
list_data <- list_settings |> 
  map(\(settings) get_data(settings$`Treat-Cov`, settings$`Cov-Outcome`))
```

## True causal effect function

Under the simulation design, the true causal effect function is

$$
\mu(t) = 2\sqrt{2}\sin(2\pi t) + \sqrt{2}\cos(2\pi t) + 
\frac{\sqrt{2}}{2}\sin(4\pi t) + \frac{\sqrt{2}}{2}\cos(4\pi t).
$$

Using the eigenfunctions defined in the simulation,

$$
\phi_{2j-1}(t)=\sqrt{2}\sin(2\pi jt),
\qquad
\phi_{2j}(t)=\sqrt{2}\cos(2\pi jt),
$$

this can equivalently be written as

$$
\mu(t) = 2\phi_1(t) + \phi_2(t) + 0.5\phi_3(t) + 0.5\phi_4(t).
$$

This equivalent representation is used below to evaluate the true effect
function on the simulation time grid.

```{r true-effect}
dummy_data <- list_data[[1]][[1]]
true_effects_mu <- drop(t(dummy_data$true_eigenfunction_phi_values) %*% 
                          data_generation$true_effect_coefficients)
```

## Evaluation criteria

For each replication, the integrated squared error (ISE) is computed as

$$
\mathrm{ISE} = \int_0^1 \left\{ \widehat{\mu}(t)-\mu(t) \right\}^2 dt.
$$

The integral is approximated numerically using the trapezoidal rule.

Across the 200 replications, we report the median ISE (MISE), the average
ISE (AISE), and the integrated squared bias (ISB),

$$
\mathrm{ISB} = \int_0^1 \left\{ \overline{\widehat{\mu}}(t)-\mu(t) \right\}^2 dt,
$$

where $\overline{\hat{\mu}}(t)$ denotes the average estimated causal
effect function across replications.

## Estimation

For each simulated data set, we estimate the causal effect function using
either the unweighted estimator or the SFPS-weighted estimator proposed
in the paper. The unweighted estimator is included as a baseline for
comparison.

For the weighted estimator, the truncation level $L$ used to estimate the
SFPS weights is selected according to the PVE threshold specified by
`PVE`. Separately, the truncation level $L^*$ used in the outcome
regression is selected according to the PVE threshold specified by
`PVE*`.

Thus, `PVE` determines the number of FPC scores used to estimate the SFPS
weights, whereas `PVE*` determines the number of FPC scores used to
estimate the causal effect function.

```{r simulation-function}
n_time_points <- length(dummy_data$time_grid)

dx <- diff(dummy_data$time_grid)
my_integrate <- function(x) {
  sum(dx * (x[-1] + x[-length(x)]) / 2)
}

run_simulation <- function(data_list, settings) {
  estimated_effects <- matrix(nrow = n_replications, ncol = n_time_points)
  ise <- double(n_replications)
  for (j in seq_len(n_replications)) {
    data <- data_list[[j]]
    
    pve <- settings$PVE
    pve_star <- settings$`PVE*`
    use_weight <- settings$Weight == "Weighted"
    
    mu_hat <- estimate$estimate_causal_effect(
      data$outcome_Y, data$functional_treatment_X, data$time_grid,
      data$covariates_C, pve, pve_star, use_weight)
    
    estimated_effects[j, ] <- mu_hat
    
    squared_error <- (mu_hat - true_effects_mu)^2
    ise[j] <- my_integrate(squared_error)
  }
  
  isb <- my_integrate((colMeans(estimated_effects) - true_effects_mu)^2)
  tibble(MISE = median(ise), AISE = mean(ise), ISB = isb)
}
```

## Monte Carlo simulation

For each simulation setting, the estimator is applied to the 200 data sets
generated above. The simulation settings are evaluated in parallel for
computational efficiency, and the resulting effect estimates are summarized
using MISE, AISE, and ISB.

```{r run-simulation, cache=TRUE}
library(future)
library(furrr)

n_workers <- min(availableCores(), nrow(simulation_settings))
plan(multisession, workers = n_workers)

result <- future_map2_dfr(list_data, list_settings, run_simulation)

df <- cbind(simulation_settings, result)
```

## Results

The results are presented separately according to the PVE threshold used
to select $L^*$ in the outcome regression.

The `Weight` column indicates whether the proposed SFPS weighting method
is used; `Unweighted` denotes the baseline estimator without SFPS weighting.

### $PVE_{L^*} = 0.95$

```{r table1-0.95}
knitr::kable(df |> filter(`PVE*` == "0.95"), digits = 4)
```

### $PVE_{L^*} = 0.99$

```{r table1-0.99}
knitr::kable(df |> filter(`PVE*` == "0.99"), digits = 4)
```
