---
title: "Reproduction of Table 1"
output: github_document
---

<!-- Table1.md is generated from Table1.Rmd Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
```

This document reproduces the simulation study reported in Table 1 of the
paper.

> Ciardulli, S., Fontana, N., Vantini, S., & Ieva, F. (2026). Generalized
> propensity score weighting for functional causal inference framework.
> https://arxiv.org/abs/2608.03200

## Simulation data

For each combination of the treatment–covariate and covariate–outcome
relationships, we generate 200 independent data sets with a sample size
of 200.
The same simulated data sets are reused across the different weighting
and PVE specifications within each data-generating scenario.

```{r settings}
library(tidyverse)
box::use(./R/data_generation)
box::use(estimate = ./R/estimate_causal_effect)

simulation_settings <- tribble(
  ~ `Treat-Cov`, ~ `Cov-Outcome`, ~ `PVE`, ~ `PVE*`, ~ `Weight`, 
    "Linear",      "Linear",        "",      "0.95",   "Unweighted", 
    "Linear",      "Linear",        "",      "0.99",   "Unweighted", 
    "Linear",      "Linear",        "0.95",  "0.95",   "Weighted",
    "Linear",      "Linear",        "0.95",  "0.99",   "Weighted",
    "Linear",      "Linear",        "0.99",  "0.95",   "Weighted",
    "Linear",      "Linear",        "0.99",  "0.99",   "Weighted",
    "Non-linear",  "Linear",        "",      "0.95",   "Unweighted", 
    "Non-linear",  "Linear",        "",      "0.99",   "Unweighted", 
    "Non-linear",  "Linear",        "0.95",  "0.95",   "Weighted",
    "Non-linear",  "Linear",        "0.95",  "0.99",   "Weighted",
    "Non-linear",  "Linear",        "0.99",  "0.95",   "Weighted",
    "Non-linear",  "Linear",        "0.99",  "0.99",   "Weighted",
    "Linear",      "Non-linear",    "",      "0.95",   "Unweighted", 
    "Linear",      "Non-linear",    "",      "0.99",   "Unweighted", 
    "Linear",      "Non-linear",    "0.95",  "0.95",   "Weighted",
    "Linear",      "Non-linear",    "0.95",  "0.99",   "Weighted",
    "Linear",      "Non-linear",    "0.99",  "0.95",   "Weighted",
    "Linear",      "Non-linear",    "0.99",  "0.99",   "Weighted",
    "Non-linear",  "Non-linear",    "",      "0.95",   "Unweighted", 
    "Non-linear",  "Non-linear",    "",      "0.99",   "Unweighted", 
    "Non-linear",  "Non-linear",    "0.95",  "0.95",   "Weighted",
    "Non-linear",  "Non-linear",    "0.95",  "0.99",   "Weighted",
    "Non-linear",  "Non-linear",    "0.99",  "0.95",   "Weighted",
    "Non-linear",  "Non-linear",    "0.99",  "0.99",   "Weighted",
)

n_replications <- 200 # R in the paper
sample_size <- 200    # n in the paper

get_data_impl <- function(treatment_covariate_relation, 
                          covariate_outcome_relation) {
  map(
    seq_len(n_replications), 
    \(dummy) data_generation$simulate_data(
      sample_size, treatment_covariate_relation, covariate_outcome_relation)
  )
}
get_data <- memoise::memoise(get_data_impl)

set.seed(314)
list_settings <- simulation_settings |> split(seq_len(nrow(simulation_settings)))
list_data <- list_settings |> 
  map(\(settings) get_data(settings$`Treat-Cov`, settings$`Cov-Outcome`))
```

## True causal effect function

Under the simulation design, the true causal effect function is

$$
\mu(t) = 2\phi_1(t) + \phi_2(t) + 0.5\phi_3(t) + 0.5\phi_4(t).
$$

Because the eigenfunctions used to generate the functional treatment are
known, the true effect can be evaluated directly on the simulation time
grid.

```{r true-effect}
dummy_data <- list_data[[1]][[1]]
true_effects_mu <- drop(t(dummy_data$true_eigenfunction_phi_values) %*% 
                          data_generation$true_effect_coefficients)
```

## Evaluation criteria

For each replication, the integrated squared error (ISE) is computed as

$$
\mathrm{ISE} = \int_0^1 \left\{ \hat{\mu}(t)-\mu(t) \right\}^2 dt.
$$

The integral is approximated numerically using the trapezoidal rule.

Across the 200 replications, we report the median ISE (MISE), the average
ISE (AISE), and the integrated squared bias (ISB),

$$
\mathrm{ISB} = \int_0^1 \left\{ \overline{\hat{\mu}(t)}-\mu(t) \right\}^2 dt,
$$

where $\overline{\hat{\mu}(t)}$ denotes the average estimated causal
effect function across replications.

## Estimation

For each simulated data set, we estimate the causal effect function using
either the unweighted estimator or the SFPS-weighted estimator.

For the weighted estimator, the SFPS weights are first estimated using
the $PVE_L$ threshold specified by `PVE`. Functional principal component
scores used in the outcome regression are then selected separately using
$PVE_{L^*}$ as `PVE*`. Thus, `PVE` controls the dimension used for estimating
the weights, whereas `PVE*` controls the dimension used for estimating the
causal effect function.

```{r simulation-function}
n_time_points <- length(dummy_data$time_grid)

dx <- diff(dummy_data$time_grid)
my_integrate <- function(x) {
  sum(dx * (x[-1] + x[-length(x)]) / 2)
}

run_simulation <- function(data_list, settings) {
  estimated_effects <- matrix(nrow = n_replications, ncol = n_time_points)
  ise <- double(n_replications)
  for (j in seq_len(n_replications)) {
    data <- data_list[[j]]
    
    pve <- settings$PVE
    pve_star <- settings$`PVE*`
    use_weight <- settings$Weight == "Weighted"
    
    mu_hat <- estimate$estimate_causal_effect(
      data$outcome_Y, data$functional_treatment_X, data$time_grid,
      data$covariates_C, pve, pve_star, use_weight)
    
    estimated_effects[j, ] <- mu_hat
    
    squared_error <- (mu_hat - true_effects_mu)^2
    ise[j] <- my_integrate(squared_error)
  }
  
  isb <- my_integrate((colMeans(estimated_effects) - true_effects_mu)^2)
  tibble(MISE = median(ise), AISE = mean(ise), ISB = isb)
}
```

## Monte Carlo simulation

The simulation settings are evaluated in parallel. For each setting,
the estimator is applied to the 200 previously generated data sets, and
the resulting effect functions are summarized using MISE, AISE, and ISB.

```{r run-simulation, cache=TRUE}
library(future)
library(furrr)

n_workers <- min(availableCores(), nrow(simulation_settings))
plan(multisession, workers = n_workers)

result <- future_map2_dfr(list_data, list_settings, run_simulation)

df <- cbind(simulation_settings, result)
```

## Results

### $PVE_{L^*} = 0.95$

```{r table1-0.95}
knitr::kable(df |> filter(`PVE*` == "0.95"), digits = 4)
```

### $PVE_{L^*} = 0.99$

```{r table1-0.99}
knitr::kable(df |> filter(`PVE*` == "0.99"), digits = 4)
```
